---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# LanduseHanneProblem

<!-- badges: start -->

<!-- badges: end -->

The goal of LanduseHanneProblem is to generate an interesting model with the Vilhelmsborg area for this purposes

1.  Sets:
  
  - $Cells,$: Vertex set of the spatial graph. LandusesLanduses: Name of possible land uses.

  - $E$: Edges, conections between cells
  - $Landuses$: Potential landuses of each cell
  
2. Parameters:
  
  - $Richness_{c,l}$: Number of species in each cell $c$ for each land use $l$.
  
  - $PhyloDiversity_{c,l}$: Phylogenetic diversity in each cell $c$ for each land use $l$.
  
  - $TransitionCost{c,l}$: Cost of transforming a cell $c$ to land use $l$.
  
  - $SpatialContiguityBonus$: Bonus for adjacent cells with the same landuse.
  
  - $b$: Budget.

3. Decision Variables:

  - $LanduseDecision_{c, l}$: Binary decision variable indicating whether land use $l$ is chosen for cell $c$.

4. Objective Function:

$\begin{align*}
\text{Maximize } & \text{ConservationIndex} \\
& = \sum_{l \in \text{Landuses}} \sum_{c \in \text{Cells}} \text{LanduseDecision}_{l,c} \cdot \text{Richness}_{l,c} \cdot \text{PhyloDiversity}_{l,c} \\
& \quad + \text{SpatialContiguityBonus} \cdot \sum_{(i,j) \in E} \sum_{l \in \text{Landuses}} \text{LanduseDecision}_{l,i} \cdot \text{LanduseDecision}_{l,j}
\end{align*}$

Constraints:

  - Proportional Use Constraint:
  
$\text{Subject to ProportionalUse}_{c}: \sum_{l \in \text{Landuses}} \text{LanduseDecision}_{l,c} \cdot \text{SuitabilityLanduse}_{l,c} = 1 \quad \forall c \in \text{Cells}$

  - Budget Constraint:

$\text{Subject to Budget}: \sum_{l \in \text{Landuses}} \sum_{c \in \text{Cells}} \text{LanduseDecision}_{l,c} \cdot \text{TransitionCost}_{l,c} = b$
    
  - Minimum Cells per landuse Contraint:
    
$\text{Subject to MinimumCellsPerLandUse}_{l}: \sum_{c \in \text{Cells}} \text{LanduseDecision}_{l, c} \geq 1 \quad \forall l \in \text{Landuses}$

Here, $LanduseDecision_{c, l}$ is a binary decision variable that is 1 if land use $l$ is chosen for cell $c$ and 0 otherwise. The objective function aims to maximize the conservation index, which is the product of richness, phylogenetic diversity, and the decision variable, which has a bonus if adjacent cells are of the same landuse.

The Proportional Use constraint ensures that for each cell, the sum of the decision variables weighted by suitability is equal to 1,indicating that exactly one land use is chosen for each cell.

The Budget constraint ensures that the total cost of transitioning cells to their chosen land uses does not exceed the budget $b$

Minimum Cells per landuse Contraint ensures that we at least have a certain number of each landuse

This model can be solved using linear programming techniques to find the optimal land-use decisions that maximize the conservation index within the given budget.

All this is summarized in this code for ampl

```{bash, eval = F}
set Cells;   # vertex set of the spacial graph
set Landuses; # Name of possible Landuses
set E within {Cells,Cells};   # Connections between cells

param Richness {Landuses, Cells}; # Number of species in each cell for each landuse

param PhyloDiversity {Landuses,Cells}; # Phylogenetic diversity in each cell for each landuse

param TransitionCost {Landuses, Cells}; # Cost of transforming a cell to this landuse

param b; #budget

param SpatialContiguityBonus; # Bonus for adjacent cells with the same landuse

var LanduseDecision {l in Landuses, c in Cells} binary; # decision on which landuse to use for cell

maximize ConservationIndex:
  sum{l in Landuses, c in Cells} LanduseDecision[l,c]*Richness[l,c]*PhyloDiversity[l,c] +
  SpatialContiguityBonus * sum{(i,j) in E, l in Landuses} LanduseDecision[l,i] * LanduseDecision[l,j];


subj to PropotionalUse{c in Cells}:
  sum{l in Landuses} LanduseDecision[l,c] <= 1;

# Ensure that at least x cells are selected for each land use
subj to MinimumCellPerLandUse{l in Landuses}:
  sum{c in Cells} LanduseDecision[l, c] >= 1;

subj to Budget:
  sum{l in Landuses, c in Cells} LanduseDecision[l,c]*TransitionCost[l,c] = b;
```

# Problem generation

## Loading packages

For this we will use the following R packages

```{r}
library(terra)
library(ggplot2)
library(TroublemakeR)
library(stringr)
library(tidyterra)
```


## Making the variables

### Phylogenetic diversity

First we start generating the Phylogenetic diversity stack:

```{r}
PD <- list.files(path = "Vilhelmsborg_test/", pattern = "^PD_.*\\.tif$", full.names = T) |> rast()

PDNames <- list.files(path = "Vilhelmsborg_test/", pattern = "^PD_.*\\.tif$", full.names = F) |> 
  stringr::str_remove_all(".tif") |> 
  stringr::str_remove_all("PD_")

names(PD) <- PDNames
```
 
This dataset looks like this:

```{r figure1, echo = F}
ggplot() + geom_spatraster(data =PD) + facet_wrap(~lyr) + theme_void() + scale_fill_viridis_b()
```


This is problematic, since we need that if there are some NAs in layers where other cells have values, this will generate a problem, to solve that we use the following code

```{r}
# Turn all NAs into 0
PD[is.na(PD)] <- 0
# The make a raster where you sum all the values
SumPD <- app(PD, sum) 
# Finally where everything is 0 based on SumPD, turn it back to NA
PD[SumPD == 0] <- NA
```

With this we change the layer into this figure:

```{r figure2, echo = F}
ggplot() + geom_spatraster(data =PD) + facet_wrap(~lyr) + theme_void() + scale_fill_viridis_b()
```

Now all the cells that have a value in one landuse will have a value in all landuses, but all other cells are just NA to avoid having to take more decisions than necessary.

We also need to normalize of values from 0 to 1, in order to then compare it later with Richness, as this give equal weight to both layers

```{r, Numbner of cells}
NormPD <- round((PD/max(minmax(PD))), 2)
```

This results in the following plot:

```{r figure3, echo = F}
ggplot() + geom_spatraster(data =NormPD) + facet_wrap(~lyr) + theme_void() + scale_fill_viridis_b()
```

### Species richness

We use the same code for richness

```{r Richness}
Richness <- list.files(path = "Vilhelmsborg_test/", pattern = "^richness_.*\\.tif$", full.names = T) |> rast()

RichnessNames <- list.files(path = "Vilhelmsborg_test/", pattern = "^richness_.*\\.tif$", full.names = F) |> 
  stringr::str_remove_all(".tif") |> 
  stringr::str_remove_all("richness_")

names(Richness) <- RichnessNames

# Turn all NAs into 0
Richness[is.na(Richness)] <- 0
# The make a raster where you sum all the values
SumRichness <- app(Richness, sum) 
# Finally where everything is 0 based on SumPD, turn it back to NA
Richness[SumRichness == 0] <- NA

NormRichness <- round((Richness/max(minmax(Richness))), 2)
```

This turns into the following figure:

```{r figure4, echo = F}
ggplot() + geom_spatraster(data =NormRichness) + facet_wrap(~lyr) + theme_void() + scale_fill_viridis_b()
```

## Number of cells for a desicion

```{r}
n_cell <- sum(!is.na(values(PD[[1]])))
```

the number of cells where desicions can be taken are `r n_cell`

## Generate problem


```{r, eval=F, cache = T}
# Define Cells
TroublemakeR::define_cells(Rasterdomain = NormPD[[1]], name = "Problem2")

# Define Edges
TroublemakeR::find_connections(Rasterdomain = NormPD[[1]], name = "Problem2")

TroublemakeR::landuse_names(landuses = PDNames, name = "Problem2")

TroublemakeR::species_suitability(Rastercurrent = NormPD, species_names = PDNames, parameter = "PhyloDiversity", name = "Problem2")

TroublemakeR::species_suitability(Rastercurrent = NormRichness, species_names = RichnessNames, parameter = "Richness", name = "Problem2")

TroublemakeR::write_ampl_lines("param TransitionCost default 1", name = "Problem2")

TroublemakeR::write_ampl_lines("param b := 100", name = "Problem2")

TroublemakeR::write_ampl_lines("param SpatialContiguityBonus := 0.2", name = "Problem2")
```


```{r}
library(readr)
library(dplyr)
Solution2 <- read_table("Solution2.txt", 
    col_names = FALSE) |> magrittr::set_colnames(c("cell", "landuse", "des")) |> dplyr::filter(des == 1)

Solution1 <- read_table("Solution1.txt", 
    col_names = FALSE) |> magrittr::set_colnames(c("cell", "landuse", "des")) |> dplyr::filter(des == 1)
```


```{r}
Temp2 <- PD[[1]]
values(Temp2) <- NA
values(Temp2)[Solution2$cell] <- Solution2$landuse

Temp1 <- PD[[1]]
values(Temp1) <- NA
values(Temp1)[Solution1$cell] <- Solution1$landuse
```

```{r}
Both <- c(Temp1, Temp2)
names(Both) <- c("Bonus_0", "Bonus_0.2")
ggplot() + geom_spatraster(data = Both) + facet_wrap(~lyr) + theme_void()
```

